---
title: "Group assignment on allocation on people into group to obtain equally efficient groups"
output: html_notebook
---

```{r}
library(lpSolveAPI)
x<- read.lp("Grp_Assignment.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
```


title: "Goal programming 1"

```{r}
library(lpSolveAPI)
x<- read.lp("Goal_Pgm_1.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
```
Result:833 employees above the goal

title: "Goal_Pgm_1"

```{r}
library(lpSolveAPI)
x<- read.lp("Goal_Pgm_2.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
get.basis(x)
```

```{r}
library(lpSolveAPI)
x<- read.lp("Goal_Pgm_3.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
get.basis(x)
```

```{r}
library(lpSolveAPI)
x<- read.lp("Goal_Pgm_4.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
get.basis(x)
```
check $duals for finding peers for branch 2. Branch 1 and branch 3.Branch 2 should move 15% towards branch 1....

```{r}
library(lpSolveAPI)
x<- read.lp("DEA_1.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
get.basis(x)
```
title: "DEA"
```{r}
library(Benchmarking)
x<- matrix(c(100,100,100,100,100),ncol=1) #define inputs
y<- matrix(c(10,15,20,23,30,31,25,30,23,20),ncol=2) #define outputs
x
y
e_crs<- dea(x,y,RTS="crs",ORIENTATION = "in")
eff(e_crs) # select effieciency scores from the results
#constant return to scale
```

```{r}
peers(e_crs)
lambda(e_crs)
```
```{r}
library(Benchmarking)
x<- matrix(c(150,400,320,520,350,320,0.2,0.7,1.2,2,1.2,0.7),ncol=2) #define inputs
y<- matrix(c(14000,14000,42000,28000,19000,14000,3500,21000,10500,42000,),ncol=2) #define outputs
x
y
e_crs<- dea(x,y,RTS="crs",ORIENTATION = "in")
eff(e_crs) # select effieciency scores from the results
#constant return to scale
```

ASSIGNMENT 3 Divya
/* x1,x2,x3,x4 are variables defining quantity of raw materials bought which are less than 200*/
/* y1,y2,y3,y4 are variables defining quantity of raw materials, which is the amount above 200 on which company is supposed to get a discount*/
/* z1,z2,z3,z4 are the quantity of items which are left in inventory for usage in next month */
/* b1,b2,b3,b4 are binary variables which is 0 if quantity bought in a particular month is less than 200 and 1 if vice versa */

```{r}
library(lpSolveAPI)
x<- read.lp("Network_op_1.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
```
```{r}
library(lpSolveAPI)
x<- read.lp("Multi_objective.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
```
```{r}
library(lpSolveAPI)
x<- read.lp("Nutrition_Problem.lp")
x
solve(x)
get.objective(x)
get.variables(x)
get.constraints(x)
get.sensitivity.objex(x)
get.sensitivity.rhs(x)
```

