---
title: "R Notebook"
output: html_notebook
---
 
```{r}
#install.packages("forecast")
library(forecast)
set.seed(2018)
```

```{r}
Sept11.data <- read.csv("D:\\UniversityData\\Allstdydata\\Special topics\\Sept11Travel.csv")
Sept11.data
```
# create time series object using ts()
# ts() takes three arguments: start, end, and freq. 
# with monthly data, the frequency of periods per season is 12 (per year). 
# arguments start and end are (season number, period number) pVehicles. 
# here start is Jan 1990: start = c(1990, 1); end is Mar 2004: end = c(2004, 4).
```{r}
AirTravel.ts <- ts(Sept11.data$Air.travel, start = c(1990, 1), end = c(2004, 4), freq = 12)
AirTravel.ts
RailTravel.ts <- ts(Sept11.data$Rail.travel, start = c(1990, 1), end = c(2004, 4), freq = 12)
RailTravel.ts
VehicleTravel.ts <- ts(Sept11.data$Vehicle.travel,start = c(1990, 1), end = c(2004, 4), freq = 12)
VehicleTravel.ts
plot(AirTravel.ts, xlab = "Time", ylab = "Air Travel (in 000s)", ylim = c(29672400, 69003600))
plot(RailTravel.ts, xlab = "Time", ylab = "Rail Travel (PM)", ylim = c(326874250, 664013880))
plot(VehicleTravel.ts, xlab = "Time", ylab = "Vehicle Travel (in billions)", ylim = c(155, 262))

```
# create short time series
# use window() to create a new, shorter time series of AirTravel.ts
# for the new three-year series, start time is oct 1st 2001 and end time is Dec 1999
```{r}
AirTravel.ts.3yrs <- window(AirTravel.ts, start = c(2001, 1), end = c(2003, 9))
AirTravel.ts.3yrs
plot(AirTravel.ts.3yrs, xlab = "Time", ylab = "Air Travel (in 000s)")

RailTravel.ts.3yrs <- window(RailTravel.ts, start = c(2001, 1), end = c(2002, 10))
RailTravel.ts.3yrs
plot(RailTravel.ts.3yrs, xlab = "Time", ylab = "Air Travel (in 000s)")

VehicleTravel.ts.3yrs <- window(VehicleTravel.ts, start = c(2001, 1), end = c(2002, 10))
VehicleTravel.ts.3yrs
plot(VehicleTravel.ts.3yrs, xlab = "Time", ylab = "Air Travel (in 000s)")
```
# fit a linear regression model to the time series
```{r}
AirTravel.lm <- forecast::tslm(AirTravel.ts ~ trend + I(trend^2))

```
# shorter and longer time series

```{r}
par(mfrow = c(2, 1))
plot(AirTravel.ts, xlab = "Time", ylab = "AirTravel (in 000s)")
#plot(AirTravel.ts, xlab = "Time", ylab = "Ridership (in 000s)", ylim = c(1300, 2300))
# overlay the fitted values of the linear model
lines(AirTravel.lm$fitted, lwd = 1)

```
Now we do Naive forecasts
First split the data into training and validation
```{r}
nValid <- 31
nTrain <- length(AirTravel.ts) - nValid
# partition the data
AirTrain.ts <- window(AirTravel.ts, start = c(1990, 1), end = c(1990, nTrain))
AirValid.ts <- window(AirTravel.ts, start = c(1990, nTrain + 1),
                   end = c(1990, nTrain + nValid))
AirTrain.ts
AirValid.ts
 

#nValid <- 31
nTrain <- length(RailTravel.ts) - nValid
# partition the data
RailTrain.ts <- window(RailTravel.ts, start = c(1990, 1), end = c(1990, nTrain))
RailValid.ts <- window(RailTravel.ts, start = c(1990, nTrain + 1),
                   end = c(1990, nTrain + nValid))
RailTrain.ts
RailValid.ts

#nValid <- 31
nTrain <- length(VehicleTravel.ts) - nValid
# partition the data
VehicleTrain.ts <- window(VehicleTravel.ts, start = c(1990, 1), end = c(1990, nTrain))
VehicleValid.ts <- window(VehicleTravel.ts, start = c(1990, nTrain + 1),end = c(1990, nTrain + nValid))
VehicleTrain.ts
VehicleValid.ts
```

```{r}
#  generate the naive and seasonal naive forecasts
Airnaive.pred <- forecast::naive(AirTrain.ts, h = nValid)
Airsnaive.pred <- forecast::snaive(AirTrain.ts, h = nValid)
Airnaive.pred
Airsnaive.pred

Railnaive.pred <- forecast::naive(RailTrain.ts, h = nValid)
Railsnaive.pred <- forecast::snaive(RailTrain.ts, h = nValid)
Railnaive.pred
Railsnaive.pred

Vehiclenaive.pred <- forecast::naive(VehicleTrain.ts, h = nValid)
Vehiclesnaive.pred <- forecast::snaive(VehicleTrain.ts, h = nValid)
Vehiclenaive.pred
Vehiclesnaive.pred
```

```{r}
# plot forecasts and actuals in the training and validation sets
plot(AirTrain.ts,  ylab = "Air Travel", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "") 
axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
lines(Airnaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(Airsnaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(AirValid.ts, col = "grey20", lty = 3)
lines(c(2004.70 - 3, 2004.70 - 3), c(0, 700999930)) 
lines(c(2004.25, 2004.25), c(0, 700999930))
text(1994.25, 69103630, "Training")
text(2002.75, 69999990, "Validation")
text(2004.90, 69103630, "Future")
arrows(2001, 67000000, 1990, 67000000, code = 3, length = 0.1, lwd = 1,angle = 30)


plot(RailTrain.ts,  ylab = "Rail Travel", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "") 
axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
lines(Railnaive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(Railsnaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(RailValid.ts, col = "grey20", lty = 3)
lines(c(2004.70 - 3, 2004.70 - 3), c(0, 804013874)) 
lines(c(2004.25, 2004.25), c(0, 804013874))
text(1996, 664013874, "Training")
text(2002.75, 664013874, "Validation")
text(2004.90, 664013874, "Future")
#arrows(2001, 630000000, 1990, 630000000, code = 3, length = 0.1, lwd = 1,angle = 30)
#arrows(2004.5 - 3, 2450, 2004, 2450, code = 3, length = 0.1, lwd = 1,angle = 30)
#arrows(2004.5, 2450, 2006, 2450, code = 3, length = 0.1, lwd = 1, angle = 30)
  
  plot(VehicleTrain.ts,  ylab = "Vehicle Travel", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2005), main = "") 
  axis(1, at = seq(1990, 2005, 1), labels = format(seq(1990, 2005, 1)))
  lines(Vehiclenaive.pred$mean, lwd = 2, col = "blue", lty = 1)
  lines(Vehiclesnaive.pred$mean, lwd = 2, col = "red", lty = 1)
  lines(VehicleValid.ts, col = "grey20", lty = 3)
  lines(c(2004.70 - 3, 2004.70 - 3), c(0, 350)) 
  lines(c(2004.25, 2004.25), c(0, 270))
  text(1994.25, 240, "Training")
  text(2002.75, 255, "Validation")
  text(2004.90, 240, "Future")
arrows(2001, 67000000, 1990, 67000000, code = 3, length = 0.1, lwd = 1,angle = 30)

```

```{r}
forecast::accuracy(Airnaive.pred, AirValid.ts)
forecast::accuracy(Airsnaive.pred, AirValid.ts)

forecast::accuracy(Railnaive.pred, RailValid.ts)
forecast::accuracy(Railsnaive.pred, RailValid.ts)

forecast::accuracy(Vehiclenaive.pred, VehicleValid.ts)
forecast::accuracy(Vehiclesnaive.pred, VehicleValid.ts)

```

Now, let us model the time series using regression models
Linear Trend

```{r}
library(forecast)
Sept11.data <- read.csv("D:\\UniversityData\\Allstdydata\\Special topics\\Sept11Travel.csv")


# create time series # ts function creates trend variable 
AirTravel.ts <- ts(Sept11.data$Air.travel, start = c(1990,1), end = c(2001,10), freq = 12)
RailTravel.ts <- ts(Sept11.data$Rail.travel, start = c(1990,1), end = c(2004,10), freq = 12)
VehicleTravel.ts <- ts(Sept11.data$Vehicle.travel, start = c(1990,1), end = c(2004,10), freq = 12)
# produce linear trend model
AirTravel.lm <- tslm(AirTravel.ts ~ trend)
RailTravel.lm <- tslm(RailTravel.ts ~ trend)
VehicleTravel.lm <- tslm(VehicleTravel.ts ~ trend)

# plot the series
plot(AirTravel.ts, xlab = "Time", ylab = "Air Travel", ylim = c(29672427,69003617),
     bty = "l")

lines(AirTravel.lm$fitted.values, lwd = 2)

plot(RailTravel.ts, xlab = "Time", ylab = "Rail Travel", ylim = c(326874247,664013874),
     bty = "l")
lines(RailTravel.lm$fitted.values, lwd = 2)

plot(VehicleTravel.ts, xlab = "Time", ylab = "Vehicle Travel", ylim = c(153.2544,261.3000),
     bty = "l")
lines(VehicleTravel.lm$fitted.values, lwd = 2)

```

Now, run the model using validation and training sets
```{r}
# fit linear trend model to training set and create forecasts
Airtrain.lm <- tslm(AirTrain.ts ~ trend)
Airtrain.lm.pred <- forecast(Airtrain.lm, h = nValid, level = 0)
par(mfrow = c(2, 1))
plot(Airtrain.lm.pred, ylim = c(29672427,69003617),  ylab = "AirTrain", xlab = "Time", 
     bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(Airtrain.lm.pred$fitted, lwd = 2, col = "blue")
lines(AirValid.ts)
plot(Airtrain.lm.pred$residuals, ylab = "Forecast Errors", 
    xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(AirValid.ts - Airtrain.lm.pred$mean, lwd = 1)



Railtrain.lm <- tslm(RailTrain.ts ~ trend)
Railtrain.lm.pred <- forecast(Railtrain.lm, h = nValid, level = 0)
par(mfrow = c(2, 1))
plot(Railtrain.lm.pred, ylim = c(326874247,664013874),  ylab = "RailTrain", xlab = "Time", 
     bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(Railtrain.lm.pred$fitted, lwd = 2, col = "blue")
lines(RailValid.ts)
plot(Railtrain.lm.pred$residuals, ylab = "Forecast Errors", 
    xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(RailValid.ts - Railtrain.lm.pred$mean, lwd = 1)



Vehicletrain.lm <- tslm(VehicleTrain.ts ~ trend)
Vehicletrain.lm.pred <- forecast(Vehicletrain.lm, h = nValid, level = 0)
par(mfrow = c(2, 1))
plot(Vehicletrain.lm.pred, ylim = c(153.2544,261.3000),  ylab = "VehicleTrain", xlab = "Time", 
     bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(Vehicletrain.lm.pred$fitted, lwd = 2, col = "blue")
lines(VehicleValid.ts)
plot(Vehicletrain.lm.pred$residuals, ylab = "Forecast Errors", 
    xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(VehicleValid.ts - Vehicletrain.lm.pred$mean, lwd = 1)
```
Summmary
```{r}
summary(Airtrain.lm)
summary(Railtrain.lm)
summary(Vehicletrain.lm)
```

Now, let's fit an exponential trend
```{r}
# fit exponential trend using tslm() with argument lambda = 0 
Airtrain.lm.expo.trend <- tslm(AirTrain.ts ~ trend, lambda = 0)
Airtrain.lm.expo.trend.pred <- forecast(Airtrain.lm.expo.trend, h = nValid, level = 0)

# fit linear trend using tslm() with argument lambda = 1 (no transform of y)
Airtrain.lm.linear.trend <- tslm(AirTrain.ts ~ trend, lambda = 1)
Airtrain.lm.linear.trend.pred <- forecast(Airtrain.lm.linear.trend, h = nValid, level = 0)

plot(Airtrain.lm.expo.trend.pred, ylim = c(29672427,69003617),  ylab = "Air Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(Airtrain.lm.expo.trend.pred$fitted, lwd = 2, col = "blue")  # Added in 6-5
lines(Airtrain.lm.linear.trend.pred$fitted, lwd = 2, col = "black", lty = 3)
lines(Airtrain.lm.linear.trend.pred$mean, lwd = 2, col = "black", lty = 3)
lines(AirValid.ts)

# For Rail
Railtrain.lm.expo.trend <- tslm(RailTrain.ts ~ trend, lambda = 0)
Railtrain.lm.expo.trend.pred <- forecast(Railtrain.lm.expo.trend, h = nValid, level = 0)
Railtrain.lm.linear.trend <- tslm(RailTrain.ts ~ trend, lambda = 1)
Railtrain.lm.linear.trend.pred <- forecast(Railtrain.lm.linear.trend, h = nValid, level = 0)

plot(Railtrain.lm.expo.trend.pred, ylim = c(326874247,664013874),  ylab = "Rail Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(Railtrain.lm.expo.trend.pred$fitted, lwd = 2, col = "blue")  # Added in 6-5
lines(Railtrain.lm.linear.trend.pred$fitted, lwd = 2, col = "black", lty = 3)
lines(Railtrain.lm.linear.trend.pred$mean, lwd = 2, col = "black", lty = 3)
lines(RailValid.ts)

# For Vehicle
Vehicletrain.lm.expo.trend <- tslm(VehicleTrain.ts ~ trend, lambda = 0)
Vehicletrain.lm.expo.trend.pred <- forecast(Vehicletrain.lm.expo.trend, h = nValid, level = 0)
Vehicletrain.lm.linear.trend <- tslm(VehicleTrain.ts ~ trend, lambda = 1)
Vehicletrain.lm.linear.trend.pred <- forecast(Vehicletrain.lm.linear.trend, h = nValid, level = 0)

plot(Vehicletrain.lm.expo.trend.pred, ylim = c(153.2544,261.3000),  ylab = "Vehicle Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(Vehicletrain.lm.expo.trend.pred$fitted, lwd = 2, col = "blue")  # Added in 6-5
lines(Vehicletrain.lm.linear.trend.pred$fitted, lwd = 2, col = "black", lty = 3)
lines(Vehicletrain.lm.linear.trend.pred$mean, lwd = 2, col = "black", lty = 3)
lines(VehicleValid.ts)


summary(Airtrain.lm.linear.trend)
summary(Railtrain.lm.linear.trend)
summary(Vehicletrain.lm.linear.trend)
```

Now, fit a quadratic trend
```{r}
#fit quadratic trend using function I(), which treats an object "as is".
Airtrain.lm.poly.trend <- tslm(AirTrain.ts ~ trend + I(trend^2))
summary(Airtrain.lm.poly.trend)
Airtrain.lm.poly.trend.pred <- forecast(Airtrain.lm.poly.trend, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(Airtrain.lm.poly.trend.pred, ylim = c(29672427,69003617),  ylab = "Air Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1))) 
lines(Airtrain.lm.poly.trend$fitted, lwd = 2)
lines(AirValid.ts)

plot(Airtrain.lm.poly.trend$residuals,  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(AirValid.ts - Airtrain.lm.poly.trend.pred$mean, lwd = 1)


#For Rail
Railtrain.lm.poly.trend <- tslm(RailTrain.ts ~ trend + I(trend^2))
summary(Railtrain.lm.poly.trend)
Railtrain.lm.poly.trend.pred <- forecast(Railtrain.lm.poly.trend, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(Railtrain.lm.poly.trend.pred, ylim = c(326874247,664013874),  ylab = "Rail Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1))) 
lines(Railtrain.lm.poly.trend$fitted, lwd = 2)
lines(RailValid.ts)

plot(Railtrain.lm.poly.trend$residuals,  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(RailValid.ts - Railtrain.lm.poly.trend.pred$mean, lwd = 1)


#For Vehicle
Vehicletrain.lm.poly.trend <- tslm(VehicleTrain.ts ~ trend + I(trend^2))
summary(Vehicletrain.lm.poly.trend)
Vehicletrain.lm.poly.trend.pred <- forecast(Vehicletrain.lm.poly.trend, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(Vehicletrain.lm.poly.trend.pred, ylim = c(153.2544,261.3000),  ylab = "Vehicle Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2)
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1))) 
lines(Vehicletrain.lm.poly.trend$fitted, lwd = 2)
lines(VehicleValid.ts)

plot(Vehicletrain.lm.poly.trend$residuals,  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(VehicleValid.ts - Vehicletrain.lm.poly.trend.pred$mean, lwd = 1)
```

Now, let us add seasonality
Our new file has a column called Season

First split the data into training and validation
```{r}
#ridership.ts <- ts(AmtrakTS.data$Ridership, start = c(1991, 1), end = c(2004, 3), freq = 12)
#nValid <- 31
#nTrain <- length(ridership.ts) - nValid
# partition the data
#train.ts <- window(ridership.ts, start = c(1991, 1), end = c(1991, nTrain))
#valid.ts <- window(ridership.ts, start = c(1991, nTrain + 1),end = c(1991, nTrain + nValid))

```

# include season as a predictor in tslm(). Here it creates 11 dummies
```{r}
Airtrain.lm.season <- tslm(AirTrain.ts ~ season)
summary(Airtrain.lm.season)

Railtrain.lm.season <- tslm(RailTrain.ts ~ season)
summary(Railtrain.lm.season)

Vehicletrain.lm.season <- tslm(VehicleTrain.ts ~ season)
summary(Vehicletrain.lm.season)
```

Now add trend to the season data
```{r}
#The graph has constant seasonal variation,as it does not change when TS increase, so We are using Additive TS instead of Multiplicative TS.
Airtrain.lm.trend.season <- tslm(AirTrain.ts ~ trend + I(trend^2) + season)
summary(Airtrain.lm.trend.season)
Airtrain.lm.trend.season.pred <- forecast(Airtrain.lm.trend.season, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(Airtrain.lm.trend.season.pred, ylim = c(29672427,69003617),  ylab = "Air Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2,col="red")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1))) 
lines(Airtrain.lm.trend.season$fitted, lwd = 2)
lines(AirValid.ts)

plot(Airtrain.lm.trend.season$residuals,  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(AirValid.ts - Airtrain.lm.trend.season.pred$mean, lwd = 1)

#For Rail
Railtrain.lm.trend.season <- tslm(RailTrain.ts ~ trend + I(trend^2) + season)
summary(Railtrain.lm.trend.season)
Railtrain.lm.trend.season.pred <- forecast(Railtrain.lm.trend.season, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(Railtrain.lm.trend.season.pred, ylim = c(326874247,664013874),  ylab = "Rail Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2,col="red")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1))) 
lines(Railtrain.lm.trend.season$fitted, lwd = 2)
lines(RailValid.ts)

plot(Railtrain.lm.trend.season$residuals,  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(RailValid.ts - Railtrain.lm.trend.season.pred$mean, lwd = 1)

#For Vehicle
Vehicletrain.lm.trend.season <- tslm(VehicleTrain.ts ~ trend + I(trend^2) + season)
summary(Vehicletrain.lm.trend.season)
Vehicletrain.lm.trend.season.pred <- forecast(Vehicletrain.lm.trend.season, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(Vehicletrain.lm.trend.season.pred, ylim = c(153.2544,261.3000),  ylab = "Vehicle Travel", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "", flty = 2,col="red")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1))) 
lines(Vehicletrain.lm.trend.season$fitted, lwd = 2)
lines(VehicleValid.ts)

plot(Vehicletrain.lm.trend.season$residuals,  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1990,2006.25), main = "")
axis(1, at = seq(1990, 2006, 1), labels = format(seq(1990, 2006, 1)))
lines(VehicleValid.ts - Vehicletrain.lm.trend.season.pred$mean, lwd = 1)
```
```{r}
# Using STL and Decompose function to suppress Seasonality
require(graphics)
DecomposeAirTrain.ts<-ts(Sept11.data$Air.travel,start = c(1998,1),end = c(2001,10),freq = 4)
Decompose_Air=decompose(DecomposeAirTrain.ts,"additive")
plot(as.ts(Decompose_Air$seasonal))
plot(as.ts(Decompose_Air$trend))
plot(as.ts(Decompose_Air$random))
plot(Decompose_Air)
air<-DecomposeAirTrain.ts-Decompose_Air$seasonal
plot(as.ts(air))
```
```{r}
#For Rail
DecomposeRailTrain.ts<-ts(Sept11.data$Rail.travel,start = c(1998,1),end = c(2001,10),frequency = 4)
Decompose_Rail=decompose(DecomposeRailTrain.ts,"additive")
plot(as.ts(Decompose_Rail$seasonal))
plot(as.ts(Decompose_Rail$trend))
plot(as.ts(Decompose_Rail$random))
plot(Decompose_Rail)
Rail<-DecomposeRailTrain.ts-Decompose_Rail$seasonal
plot(as.ts(Rail))
```
```{r}
#For Vehicle
DecomposeVehicleTrain.ts<-ts(Sept11.data$Vehicle.travel,start = c(2000,1),end = c(2003,10),frequency = 4)
Decompose_Vehicle=decompose(DecomposeVehicleTrain.ts,"additive")
plot(as.ts(Decompose_Vehicle$seasonal))
plot(as.ts(Decompose_Vehicle$trend))
plot(as.ts(Decompose_Vehicle$random))
plot(Decompose_Vehicle)
Vehicle<-DecomposeVehicleTrain.ts-Decompose_Vehicle$seasonal
plot(as.ts(Vehicle))
```

```{r}

#Checking whether our Time series is staionary 
#Time series is stationary if its mean level and variance stay steady over time
library(lmtest)
if (gqtest(AirTrain.ts~1)$p.value < 0.10) {
    x<-log(AirTrain.ts)
}

 Acf(x)
 Pacf(x)
#check if the lags die out
 
Box.test(x,lag=25,type="Ljung-Box")
#p value <0.05 suggest that our TS is staionary
```





