---
title: "Simulated Annealing - GenSA"
output: html_notebook
Comaprison of different algos
---
References: 
1. https://journal.r-project.org/archive/2013/RJ-2013-002/RJ-2013-002.pdf
2. https://www.jstatsoft.org/index.php/jss/article/view/v060i06/v60i06.pdf
3. http://toddwschneider.com/posts/traveling-salesman-with-simulated-annealing-r-and-shiny/

Plot Acceptance Probability for SA
```{r}
#install packages ggplot and colorspace
library(ggplot2)
ap <- function(x) {
  exp((2-x)/5)
}
pl <- ggplot(data.frame(x=c(2,7)),aes(x=x))+stat_function(fun=ap)
pl
```

Install package 
```{r}
#install.packages("GenSA")
```
Define the test example
```{r}
fn.call <- 0
Rastrigin <- function(x) {
  fn.call <<- fn.call + 1
  sum(x^2 - 10 * cos(2 * pi * x)) + 10 * length(x)
}
```

solving using Optim and then GenSA
```{r}
set.seed(1234)    # setting random  initially
options(digits = 10)
dimension <- 2     # maximum is 30
lower <- rep(-5.12, dimension)
upper <- rep(5.12, dimension)
x.ini <- lower + runif(length(lower)) * (upper - lower)
fn.call <- 0
out.sann <- optim(par = x.ini, fn = Rastrigin, method = "SANN")    #simulated annealing
out.sann[c("value","par","counts")]
```
```{r}
set.seed(1234)
library(GenSA)
expected.val <- 0
absTol <- 1e-13    #tolerance we set,how close you want answer,slight larger than machine error
fn.call <- 0
out.GenSA <- GenSA(par = NULL, lower = lower, upper = upper, fn = Rastrigin, control = list(threshold.stop = expected.val + absTol))
out.GenSA[c("value", "par", "counts")]
cat("GenSA call functions", fn.call, "times.\n")
```
Now, let's repeat the experiment with Dimension = 30.  (Yao et al., 1999; Mebane Jr and Sekhon, 2011)
```{r}
set.seed(1234)
dimension <- 30
lower <- rep(-5.12, dimension)
upper <- rep(5.12, dimension)
fn.call <- 0
out.GenSA <- GenSA(lower = lower, upper = upper, fn = Rastrigin, control = list(max.time=5, verbose=TRUE))
out.GenSA[c("value")]

```
Let's use SA on our book test function. Note that we are converting a
maximization problem to a minimization problem

```{r}
testfun <- function(x) {
  -(12*x^5-975*x^4+28000*x^3-345000*x^2+1800000*x)
}
library(ggplot2)
p9 <- ggplot(data.frame(x = c(0, 30)), aes(x = x)) + stat_function(fun = testfun)
p9
set.seed(1234)
lower <- 0
upper <- 30
out.GenSA <- GenSA(lower=lower, upper=upper, fn = testfun)
out.GenSA[c("value","par")]   # answer is location x=20 and global minimum is -44000000..

```
Risk allocation portfolios

Mean-risk models were developed in the 1950s for portfolio selection problems. Value-at-Risk (VaR) and Conditional Value-at-Risk (CVaR) are the most popular measures of downside risk. Mullen et al. (2011) and Ardia et al. (2011) used DEoptim to find the portfolio weights for which the portfolio has the lowest CVaR and each investment can contribute at most 22.5% to total portfolio CVaR risk.For details, please refer to Mullen et al. (2011); Ardia et al. (2011). The code for objective function in portfolio optimization are from Ardia et al. (2011).
```{r}
library("quantmod")
tickers <- c("GE", "IBM", "JPM", "MSFT", "WMT")
getSymbols(tickers, from = "2010-12-01", to = "2016-12-31",src="google")
P <- NULL
for(ticker in tickers) {
tmp <- Cl(to.monthly(eval(parse(text = ticker))))
P <- cbind(P, tmp)
}
colnames(P) <- tickers
R <- diff(log(P))
R <- R[-1,]
mu <- colMeans(R)
sigma <- cov(R)
#install.packages("PerformanceAnalytics")
library("PerformanceAnalytics")
pContribCVaR <- ES(weights = rep(0.2, 5), method = "gaussian", portfolio_method = "component", mu = mu, sigma = sigma)$pct_contrib_ES
obj <- function(w) {
fn.call <<- fn.call + 1
if (sum(w) == 0) { w <- w + 1e-2 }
w <- w / sum(w)
CVaR <- ES(weights = w, method = "gaussian", portfolio_method = "component",mu = mu, sigma = sigma)
tmp1 <- CVaR$ES
tmp2 <- max(CVaR$pct_contrib_ES - 0.225, 0)
out <- tmp1 + 1e3 * tmp2
return(out)
}
# Now run GenSA
set.seed(1234)
fn.call <<- 0
out.GenSA <- GenSA(fn = obj, lower = rep(0, 5), upper = rep(1, 5),
control = list(smooth = FALSE, max.call = 3000))
fn.call.GenSA <- fn.call
out.GenSA$value
out.GenSA$counts
cat("GenSA call functions", fn.call.GenSA, "times.\n")
wstar.GenSA <- out.GenSA$par
wstar.GenSA <- wstar.GenSA / sum(wstar.GenSA)
rbind(tickers, round(100 * wstar.GenSA, 2))
100 * (sum(wstar.GenSA * mu) - mean(mu))
```

