---
title: "Time Series Forecasting"
output: html_notebook
---
Reference: 

```{r}
install.packages("forecast")
library(forecast)

```

```{r}
Amtrak.data <- read.csv("C:\\Users\\divya\\Documents\\GitHub\\60095_201810\\Amtrak.csv")
Amtrak.data
```
# creating time series object using ts()
# ts() takes three arguments: start, end, and freq. 
# with monthly data, the frequency of periods per season is 12 (per year). 
# arguments start and end are (season number, period number) pairs. 
# here start is Jan 1991: start = c(1991, 1); end is Mar 2004: end = c(2004, 3).
```{r}
ridership.ts <- ts(Amtrak.data$Ridership, 
    start = c(1991, 1), end = c(2004, 3), freq = 12)
plot(ridership.ts, xlab = "Time", ylab = "Ridership (in 000s)", ylim = c(1300, 2300))
```
# creating short time series
# window() to create a new, shorter time series of ridership.ts
# for the new three-year series, start time is Jan 1997 and end time is Dec 1999
```{r}
ridership.ts.3yrs <- window(ridership.ts, start = c(1997, 1), end = c(1999, 12))
```
# fit a linear regression model to the time series
```{r}
ridership.lm <- forecast::tslm(ridership.ts ~ trend + I(trend^2))

```
# shorter and longer time series

```{r}
par(mfrow = c(2, 1))
plot(ridership.ts.3yrs, xlab = "Time", ylab = "Ridership (in 000s)", 
                        ylim = c(1300, 2300))
plot(ridership.ts, xlab = "Time", ylab = "Ridership (in 000s)", ylim = c(1300, 2300))

# overlay the fitted values of the linear model
lines(ridership.lm$fitted, lwd = 1)

```
Now we do Naive forecasts
First split the data into training and validation
```{r}
nValid <- 36
nTrain <- length(ridership.ts) - nValid
# partitioning the data
train.ts <- window(ridership.ts, start = c(1991, 1), end = c(1991, nTrain))
valid.ts <- window(ridership.ts, start = c(1991, nTrain + 1),
                   end = c(1991, nTrain + nValid))
 
```

```{r}
#  generating the naive and seasonal naive forecasts
naive.pred <- forecast::naive(train.ts, h = nValid)
snaive.pred <- forecast::snaive(train.ts, h = nValid)
```

```{r}
# plotting forecasts and actuals in the training and validation sets
plot(train.ts, ylim = c(1300, 2600),  ylab = "Ridership", xlab = "Time", bty = "l", 
    xaxt = "n", xlim = c(1991,2006.25), main = "") 
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1)))
lines(naive.pred$mean, lwd = 2, col = "blue", lty = 1)
lines(snaive.pred$mean, lwd = 2, col = "red", lty = 1)
lines(valid.ts, col = "grey20", lty = 3)
lines(c(2004.25 - 3, 2004.25 - 3), c(0, 3500)) 
lines(c(2004.25, 2004.25), c(0, 3500))
text(1996.25, 2500, "Training")
text(2002.75, 2500, "Validation")
text(2005.25, 2500, "Future")
arrows(2004 - 3, 2450, 1991.25, 2450, code = 3, length = 0.1, lwd = 1,angle = 30)
arrows(2004.5 - 3, 2450, 2004, 2450, code = 3, length = 0.1, lwd = 1,angle = 30)
arrows(2004.5, 2450, 2006, 2450, code = 3, length = 0.1, lwd = 1, angle = 30)
```

```{r}
forecast::accuracy(naive.pred, valid.ts)
forecast::accuracy(snaive.pred, valid.ts)
```

Now, modelling the time series using regression models
Linear Trend

```{r}
library(forecast)

Amtrak.data <- read.csv("Amtrak.csv")

# creating time series # ts function creates trend variable 
ridership.ts <- ts(Amtrak.data$Ridership, start = c(1991,1),
                   end = c(2004,3), freq = 12)

#  linear trend model
ridership.lm <- tslm(ridership.ts ~ trend)

# plotting the series
plot(ridership.ts, xlab = "Time", ylab = "Ridership", ylim = c(1300,2300),
     bty = "l")
lines(ridership.lm$fitted, lwd = 2)
```

Now, run the model using validation and training sets
```{r}
# fitting linear trend model to training set and create forecasts
train.lm <- tslm(train.ts ~ trend)
train.lm.pred <- forecast(train.lm, h = nValid, level = 0)

par(mfrow = c(2, 1))
plot(train.lm.pred, ylim = c(1300, 2600),  ylab = "Ridership", xlab = "Time", 
     bty = "l", xaxt = "n", xlim = c(1991,2006.25), main = "", flty = 2)
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1)))
lines(train.lm.pred$fitted, lwd = 2, col = "blue")
lines(valid.ts)
plot(train.lm.pred$residuals, ylim = c(-420, 500),  ylab = "Forecast Errors", 
    xlab = "Time", bty = "l", xaxt = "n", xlim = c(1991,2006.25), main = "")
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1)))
lines(valid.ts - train.lm.pred$mean, lwd = 1)
```
Summmary
```{r}
summary(train.lm)
```

Now, let's fit an exponential trend
```{r}
# fitting exponential trend using tslm() with argument lambda = 0 
train.lm.expo.trend <- tslm(train.ts ~ trend, lambda = 0)
train.lm.expo.trend.pred <- forecast(train.lm.expo.trend, h = nValid, level = 0)

# fitting linear trend using tslm() with argument lambda = 1 (no transform of y)
train.lm.linear.trend <- tslm(train.ts ~ trend, lambda = 1)
train.lm.linear.trend.pred <- forecast(train.lm.linear.trend, h = nValid, level = 0)

plot(train.lm.expo.trend.pred, ylim = c(1300, 2600),  ylab = "Ridership", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1991,2006.25), main = "", flty = 2)
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1)))
lines(train.lm.expo.trend.pred$fitted, lwd = 2, col = "blue")  # Added in 6-5
lines(train.lm.linear.trend.pred$fitted, lwd = 2, col = "black", lty = 3)
lines(train.lm.linear.trend.pred$mean, lwd = 2, col = "black", lty = 3)
lines(valid.ts)
```

Now, fitting a quadratic trend as it usually gives the best result
```{r}
#fit quadratic trend using function I(), which treats an object "as is".
train.lm.poly.trend <- tslm(train.ts ~ trend + I(trend^2))
summary(train.lm.poly.trend)
train.lm.poly.trend.pred <- forecast(train.lm.poly.trend, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(train.lm.poly.trend.pred, ylim = c(1300, 2600),  ylab = "Ridership", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1991,2006.25), main = "", flty = 2)
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1))) 
lines(train.lm.poly.trend$fitted, lwd = 2)
lines(valid.ts)

plot(train.lm.poly.trend$residuals, ylim = c(-400, 550),  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1991,2006.25), main = "")
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1)))
lines(valid.ts - train.lm.poly.trend.pred$mean, lwd = 1)
```

Now, adding seasonality
New file AmtrakTS has a column called Season
```{r}
AmtrakTS.data <- read.csv("AmtrakTS.csv")
AmtrakTS.data

```
splitting the data into training and validation
```{r}
ridership.ts <- ts(AmtrakTS.data$Ridership, 
    start = c(1991, 1), end = c(2004, 3), freq = 12)
nValid <- 36
nTrain <- length(ridership.ts) - nValid
# partitioning the data
train.ts <- window(ridership.ts, start = c(1991, 1), end = c(1991, nTrain))
valid.ts <- window(ridership.ts, start = c(1991, nTrain + 1),
                   end = c(1991, nTrain + nValid))

```

# including season as a predictor in tslm(). Here it creates 11 dummies
```{r}
train.lm.season <- tslm(train.ts ~ season)
summary(train.lm.season)
```

Now adding trend to the season data
```{r}
train.lm.trend.season <- tslm(train.ts ~ trend + I(trend^2) + season)
summary(train.lm.trend.season)
train.lm.trend.season.pred <- forecast(train.lm.trend.season, h = nValid, level = 0)

par(mfrow = c(2,1))
plot(train.lm.trend.season.pred, ylim = c(1300, 2600),  ylab = "Ridership", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1991,2006.25), main = "", flty = 2,col="red")
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1))) 
lines(train.lm.trend.season$fitted, lwd = 2)
lines(valid.ts)

plot(train.lm.trend.season$residuals, ylim = c(-400, 550),  ylab = "Forecast Errors", 
 xlab = "Time", bty = "l", xaxt = "n", xlim = c(1991,2006.25), main = "")
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1)))
lines(valid.ts - train.lm.trend.season.pred$mean, lwd = 1)
```


